Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-04-05 paper.qmd
Updated with PBMC new layout (best)

<!--Fit the best model for author suggestion and compute error-->
```{r}
#| label: hexbin-pbmc
## Compute hexbin parameters
num_bins_x_pbmc <- 26

@@

## Compute hexbin parameters
num_bins_x_pbmc <- 18

@@

distance_df_small_edges_pbmc <- distance_pbmc |>
  filter(distance < 0.07) #benchmark_pbmc

@@

```{r}
#| echo: false
#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse a) and the \\gD{} layout selected (@fig-pbmc-mse b) by MSE plot (@fig-pbmc-mse) from the same NLDR with different hyper-parameter choice. The PBMC3k data ($n =  2622$) has three clusters in $9\\text{-}D$, where two of them being close and one separated from the others.  Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse a ($a_1 = 0.02$, $b = 3480/349 (60, 58)$) shows three-well separated clusters with big separations. On the other hand, the model fit with @fig-pbmc-mse b ($a_1 = 0.04$, $b = 1020/347 (30, 34)$) shows two close clusters, and one separated cluster. The edge connected the two clusters proves the closeness of the two clusters. Therefore, @fig-pbmc-mse b is more reasonable than @fig-pbmc-mse a. Also, the fitted models in $9\\text{-}D$ help to see some unobserved patterns of the data: (i) dense points, and (ii) non-linear clusters. Videos of the langevitour animations are available at <https://youtu.be/0cKX_HG_n0k> and <https://youtu.be/KhJvsRtaX04> respectively."
#| label: fig-model-pbmc-author-proj
#| fig-pos: H
#| fig-width: 10
#| fig-height: 15
#| out-width: 90%

>>>>

<!--Fit the best model for author suggestion and compute error-->
```{r}
#| label: hexbin-pbmc
## Compute hexbin parameters
num_bins_x_pbmc <- 22

@@

## Compute hexbin parameters
num_bins_x_pbmc <- 13

@@

distance_df_small_edges_pbmc <- distance_pbmc |>
  filter(distance < 0.1) #benchmark_pbmc

@@

```{r}
#| echo: false
#| fig-cap: "Compare the published \\gD{} layout (@fig-pbmc-mse a) and the \\gD{} layout selected (@fig-pbmc-mse b) by MSE plot (@fig-pbmc-mse) from the same NLDR with different hyper-parameter choice. The PBMC3k data ($n =  2622$) has three clusters in $9\\text{-}D$, where two of them being close and one separated from the others.  Two \\gD{} projection from a tour on $9\\text{-}D$ of the model fit with @fig-pbmc-mse a ($a_1 = 0.06$, $b = 616/86 (22, 28)$) shows three-well separated clusters with big separations. On the other hand, the model fit with @fig-pbmc-mse b ($a_1 = 0.1$, $b = 234/98 (13, 18)$) shows two close clusters, and one separated cluster. The edge connected the two clusters proves the closeness of the two clusters. Therefore, @fig-pbmc-mse b is more reasonable than @fig-pbmc-mse a. Also, the fitted models in $9\\text{-}D$ help to see some unobserved patterns of the data: (i) dense points, and (ii) non-linear clusters. Videos of the langevitour animations are available at <https://youtu.be/0cKX_HG_n0k> and <https://youtu.be/KhJvsRtaX04> respectively."
#| label: fig-model-pbmc-author-proj
#| fig-pos: H
#| fig-width: 10
#| fig-height: 15
#| out-width: 90%
